---
title: "Coreg_eda"
output: html_document
date: '2023-06-30'
editor_options: 
  chunk_output_type: console
---

#This rmd serves as a building block for the models we will make to identify the z-score, which is (elevation difference - predicted median) / predicted nmad. Refer to Cv_Rol_MedModel2.rmd for a more clean version of this, or you can head straight to the z_score.rmd

```{r, message = FALSE}
library(tidyverse)
library(raster)
library(terra)
library(viridis)
library(RColorBrewer)
library(ggplot2)
library(ggplotify)
rwb <- colorRampPalette(c("red", "white", "blue"))(50)

stor_no_outliers <- read_csv("data/stor_no_outliers.csv") #glacier only w/ outliers removed
```


#Elevation & Curve to Median
```{r}
#Making a dataset for the plots below
stor_no_out_plot <- stor_no_outliers %>% 
  group_by(elev_cut, plan_curve_bin, prof_curve_bin,max_curve_bin,slope_cut) %>% 
  mutate(med = median(elev_diff)) %>% 
  mutate(upper = med + mad(elev_diff)*2) %>% 
  mutate(lower = med - mad(elev_diff)*2) %>% 
  mutate(mad = mad(elev_diff)) 


#Making plots to see if the following variables effect the median of elevation difference

#Elevation (med appears to be affected by elev)
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = med))

#Curve plan
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = plan_curve_bin, y = med))

#Curve prof
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = prof_curve_bin, y = med))

#Curve max
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = max_curve_bin, y = med))

#Curve max
stor_no_out_plot %>% 
  ggplot() +
  geom_boxplot(aes(x = northness_cut, y = med)) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )

#Curve max
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = eastness_cut, y = med)) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )
```


# NMAD EDA

# elev_diff and slope using bins
```{r}
stor_no_out_plot %>% 
  group_by(slope_cut) %>% 
  summarize(mad_elev_diff = mad(elev_diff)) %>% 
  ungroup() %>% 
  ggplot(aes(x = slope_cut, y = mad_elev_diff)) +
  geom_point() +
  scale_y_continuous(limit = c(0,4))+
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )

#change scale so that we can look at all of them together 
```

# elev_diff and curvature using bins
```{r}
stor_no_out_plot %>% 
  group_by(plan_curve_bin) %>% 
  summarize(mad_elev_diff = mad(elev_diff)) %>% 
  ungroup() %>% 
  ggplot(aes(x = plan_curve_bin, y = mad_elev_diff)) +
  geom_point() +
  scale_y_continuous(limit = c(0,4))+
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )
stor_no_out_plot %>% 
  group_by(prof_curve_bin) %>% 
  summarize(mad_elev_diff = mad(elev_diff)) %>% 
  ungroup() %>% 
  ggplot(aes(x = prof_curve_bin, y = mad_elev_diff)) +
  geom_point() +
  scale_y_continuous(limit = c(0,4))+
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )

stor_no_out_plot %>% 
  group_by(max_curve_bin) %>% 
  summarize(mad_elev_diff = mad(elev_diff)) %>% 
  ungroup() %>% 
  ggplot(aes(x = max_curve_bin, y = mad_elev_diff)) +
  geom_point() +
  scale_y_continuous(limit = c(0,4))+
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )
```

# elevation difference and elevation cut
```{r}
stor_no_out_plot %>% 
  group_by(elev_cut) %>% 
  summarize(mad_elev_diff = mad(elev_diff)) %>% 
  ungroup() %>% 
  ggplot(aes(x = elev_cut, y = mad_elev_diff)) +
  geom_point() +
  scale_y_continuous(limit = c(0,4))+
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )

```

# aspect (northness and eastness)

```{r}
stor_no_out_plot %>% 
  group_by(northness_cut) %>% 
  summarize(mad_elev_diff = mad(elev_diff)) %>% 
  ungroup() %>% 
  ggplot(aes(x = northness_cut, y = mad_elev_diff)) +
  geom_point() +
  scale_y_continuous(limit = c(0,4))+
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
                )

stor_no_out_plot %>% 
  group_by(eastness_cut) %>% 
  summarize(mad_elev_diff = mad(elev_diff)) %>% 
  ungroup() %>% 
  ggplot(aes(x = eastness_cut, y = mad_elev_diff)) +
  geom_point() +
  scale_y_continuous(limit = c(0,4))+
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
                )
```


#Cleaning Set up
```{r}
stor_no_outliers <- stor_no_outliers %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
     mutate(lower_slope = as.numeric(gsub("\\((.*),.*\\]", "\\1", slope_cut)),
            upper_slope = as.numeric(gsub(".*,(.*)\\]", "\\1", slope_cut)),
          med_slope_cut = (lower_slope + upper_slope) / 2,
          cen_slope_med = med_slope_cut - median(med_slope_cut)) %>% 
     mutate(lower_plan = as.numeric(gsub("\\((.*),.*\\]", "\\1", plan_curve_bin)),
            upper_plan = as.numeric(gsub(".*,(.*)\\]", "\\1", plan_curve_bin)),
            med_plan_cut = (lower_plan + upper_plan) / 2,
            cen_plan_med = med_plan_cut - median(med_plan_cut)) %>% 
     mutate(lower_prof = ifelse(prof_curve_bin == "[0,0.2]",
                                0,
            as.numeric(gsub("\\((.*),.*\\]", "\\1", prof_curve_bin))),
         upper_prof = ifelse(prof_curve_bin == "[0,0.2]",
                                0.2,
            as.numeric(gsub(".*,(.*)\\]", "\\1", prof_curve_bin))),
          med_prof_cut = (lower_prof + upper_prof) / 2,
          cen_prof_med = med_prof_cut - median(med_prof_cut)) %>% 
     mutate(lower_max = as.numeric(gsub("\\((.*),.*\\]", "\\1", max_curve_bin)),
          upper_max = as.numeric(gsub(".*,(.*)\\]", "\\1", max_curve_bin)),
          med_max_cut = (lower_max + upper_max) / 2,
          cen_max_med = med_max_cut - median(med_max_cut))

summary(stor_no_outliers$lower_prof)


stor_med_mad <- stor_no_outliers %>% 
  group_by(elev_cut, slope_cut) %>% 
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff), 
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
     mutate(lower_slope = as.numeric(gsub("\\((.*),.*\\]", "\\1", slope_cut)),
            upper_slope = as.numeric(gsub(".*,(.*)\\]", "\\1", slope_cut)),
          med_slope_cut = (lower_slope + upper_slope) / 2,
          cen_slope_med = med_slope_cut - median(med_slope_cut)) 


stor_med_mad1 <- stor_no_outliers %>% 
  group_by(elev_cut, plan_curve_bin) %>% 
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff), 
         count = n()) %>% #median for the curve
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
     mutate(lower_plan = as.numeric(gsub("\\((.*),.*\\]", "\\1", plan_curve_bin)),
            upper_plan = as.numeric(gsub(".*,(.*)\\]", "\\1", plan_curve_bin)),
            med_plan_cut = (lower_plan + upper_plan) / 2, 
            cen_plan_med = med_plan_cut - median(med_plan_cut)) #The median of the values; Cut_number for all other variables except elev; When plotting, size = count


stor_med_mad2 <- stor_no_outliers %>% 
  group_by(elev_cut, prof_curve_bin) %>% 
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff), 
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
     mutate(lower_prof = ifelse(prof_curve_bin == "[0,0.2]",
                                0,
            as.numeric(gsub("\\((.*),.*\\]", "\\1", prof_curve_bin))),
         upper_prof = ifelse(prof_curve_bin == "[0,0.2]",
                                0.2,
            as.numeric(gsub(".*,(.*)\\]", "\\1", prof_curve_bin))),
          med_prof_cut = (lower_prof + upper_prof) / 2,
          cen_prof_med = med_prof_cut - median(med_prof_cut)) 



stor_med_mad3 <- stor_no_outliers %>% 
  group_by(elev_cut, max_curve_bin) %>% 
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff), 
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
     mutate(lower_max = as.numeric(gsub("\\((.*),.*\\]", "\\1", max_curve_bin)),
          upper_max = as.numeric(gsub(".*,(.*)\\]", "\\1", max_curve_bin)),
          med_max_cut = (lower_max + upper_max) / 2,
          cen_max_med = med_max_cut - median(med_max_cut))


stor_med_mad4 <- stor_no_outliers %>% 
  group_by(elev_cut, slope_cut, plan_curve_bin ) %>% 
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff), 
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
     mutate(lower_slope = as.numeric(gsub("\\((.*),.*\\]", "\\1", slope_cut)),
            upper_slope = as.numeric(gsub(".*,(.*)\\]", "\\1", slope_cut)),
          med_slope_cut = (lower_slope + upper_slope) / 2,
          cen_slope_med = med_slope_cut - median(med_slope_cut)) %>% 
     mutate(lower_plan = as.numeric(gsub("\\((.*),.*\\]", "\\1", plan_curve_bin)),
            upper_plan = as.numeric(gsub(".*,(.*)\\]", "\\1", plan_curve_bin)),
            med_plan_cut = (lower_plan + upper_plan) / 2,
            cen_plan_med = med_plan_cut - median(med_plan_cut)) 



stor_med_mad5 <- stor_no_outliers %>% 
  group_by(elev_cut, slope_cut, max_curve_bin) %>% 
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff), 
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
     mutate(lower_slope = as.numeric(gsub("\\((.*),.*\\]", "\\1", slope_cut)),
            upper_slope = as.numeric(gsub(".*,(.*)\\]", "\\1", slope_cut)),
          med_slope_cut = (lower_slope + upper_slope) / 2,
          cen_slope_med = med_slope_cut - median(med_slope_cut)) %>% 
      mutate(lower_max = as.numeric(gsub("\\((.*),.*\\]", "\\1", max_curve_bin)),
          upper_max = as.numeric(gsub(".*,(.*)\\]", "\\1", max_curve_bin)),
          med_max_cut = (lower_max + upper_max) / 2,
          cen_max_med = med_max_cut - median(med_max_cut))


stor_med_mad6 <- stor_no_outliers %>% 
  group_by(elev_cut, slope_cut, max_curve_bin, plan_curve_bin) %>% 
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff), 
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
     mutate(lower_slope = as.numeric(gsub("\\((.*),.*\\]", "\\1", slope_cut)),
            upper_slope = as.numeric(gsub(".*,(.*)\\]", "\\1", slope_cut)),
          med_slope_cut = (lower_slope + upper_slope) / 2,
          cen_slope_med = med_slope_cut - median(med_slope_cut)) %>% 
      mutate(lower_max = as.numeric(gsub("\\((.*),.*\\]", "\\1", max_curve_bin)),
          upper_max = as.numeric(gsub(".*,(.*)\\]", "\\1", max_curve_bin)),
          med_max_cut = (lower_max + upper_max) / 2,
          cen_max_med = med_max_cut - median(med_max_cut)) %>% 
  mutate(lower_plan = as.numeric(gsub("\\((.*),.*\\]", "\\1", plan_curve_bin)),
            upper_plan = as.numeric(gsub(".*,(.*)\\]", "\\1", plan_curve_bin)),
            med_plan_cut = (lower_plan + upper_plan) / 2,
            cen_plan_med = med_plan_cut - median(med_plan_cut)) 





sapply(lapply(stor_med_mad, unique), length)

            

stor_med_mad_test <- stor_no_outliers %>% 
  group_by(elev_cut, slope_cut)%>%
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff), 
         count = n()) %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - 1425)%>%
  mutate(lower_slope = as.numeric(gsub("\\((.*),.*\\]", "\\1", slope_cut)),
        upper_slope = as.numeric(gsub(".*,(.*)\\]", "\\1", slope_cut)),
          med_slope_cut = (lower_slope + upper_slope) / 2,
        cen_slope_med = med_slope_cut - 7.5)







sapply(lapply(test, unique), length)

summary(stor_med_mad$count)


#plot for northness and eastness 
stor_no_outliers %>% 
  ggplot() +
  geom_smooth(aes(x=northness, y=elev_diff))

stor_no_outliers %>% 
  ggplot() +
  geom_raster(aes(x=x, y=y, fill=aspect_degrees))#double check the hill shade #aspect_degrees, try and use rainbow colors
  #positive (light) means true, ie: east; negative (dark) is the opposite


# mutate(lower_northness = as.numeric(gsub("\\((.*),.*\\]", "\\1", northness_cut)),
  #        upper_northness = as.numeric(gsub(".*,(.*)\\]", "\\1", northness_cut)),
  #        med_northness_cut = (lower_aspect + upper_aspect) / 2) %>% 
  # mutate(cen_northness_med = med_aspect_cut - median(med_aspect_cut))

sapply(lapply(stor_med_mad, unique), length)
```

#Models
```{r}
# making model for MED (Poly)
#Third Order (from the reading that looked at RMSE)

#Slope (possibly)
med_model3 <- lm(med_elev_diff ~ polym(cen_elev_med, degree=3, raw=TRUE), data = stor_med_mad, weights = count)
summary(med_model3)

stor_med_mad$residuals3 <- residuals(med_model3)

med_model3.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, degree=3, raw=TRUE), data = stor_med_mad, weights = count)
summary(med_model3.1)
stor_med_mad$residuals3.1 <- residuals(med_model3.1)

stor_med_mad %>% 
  ggplot() +
  geom_point(aes(x = cen_slope_med, y = residuals3)) +
  geom_smooth(aes(x = cen_slope_med, y = residuals3)) #Slope is not significant and important to be added. Check out Northness, Eastness, plan_curve_bin, prof_curve_bin, max_curve_bin...maybe aspect_cut if needed.

#Partial F-test (Nested Models)
anova(med_model3, med_model3.1)


#PLan_Curve (f-test says we should add)
med_model4 <- lm(med_elev_diff ~ polym(cen_elev_med, degree=3, raw=TRUE), data = stor_med_mad1, weights = count)
summary(med_model4)
residuals4 <- residuals(med_model4)

med_model4.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_plan_med, degree=3, raw=TRUE), data = stor_med_mad1, weights = count)
summary(med_model4.1)

stor_med_mad1 %>% 
  ggplot() +
  geom_point(aes(x = cen_plan_med, y = residuals4)) +
  geom_smooth(aes(x = cen_plan_med, y = residuals4))

anova(med_model4, med_model4.1)

#Prof_curve (Not significant)
med_model5 <- lm(med_elev_diff ~ polym(cen_elev_med, degree=3, raw=TRUE), data = stor_med_mad2, weights = count)
summary(med_model5)
residuals5 <- residuals(med_model5)

med_model5.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_prof_med, degree=3, raw=TRUE), data = stor_med_mad2, weights = count)
summary(med_model5.1)

anova(med_model5, med_model5.1)

stor_med_mad2 %>% 
  ggplot() +
  geom_point(aes(x = cen_prof_med, y = residuals5)) +
  geom_smooth(aes(x = cen_prof_med, y = residuals5))



#Max_curve 
med_model6 <- lm(med_elev_diff ~ polym(cen_elev_med, degree=3, raw=TRUE), data = stor_med_mad3, weights = count)
summary(med_model6)
residuals6 <- residuals(med_model6)

med_model6.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_max_med,degree=3, raw=TRUE), data = stor_med_mad3, weights = count)
summary(med_model6.1)
residuals6.1 <- residuals(med_model6.1)

anova(med_model6, med_model6.1)

stor_med_mad3 %>% 
  ggplot() +
  geom_point(aes(x = cen_max_med, y = residuals6)) +
  geom_smooth(aes(x = cen_max_med, y = residuals6))

#Slope_cut & PLan_curve
med_model3.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, degree=3, raw=TRUE), data = stor_med_mad4, weights = count)
summary(med_model3.1)
stor_med_mad4$residuals3.1 <- residuals(med_model3.1)

med_model3.2 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, cen_plan_med, degree=3, raw=TRUE), data = stor_med_mad4, weights = count)
summary(med_model3.2)
stor_med_mad4$residuals3.2 <- residuals(med_model3.2)

stor_med_mad4 %>% 
  ggplot() +
  geom_point(aes(x = cen_plan_med, y = residuals3.1)) +
  geom_smooth(aes(x = cen_plan_med, y = residuals3.1))

anova(med_model3.1, med_model3.2)

#Max Curve and Slope cut

med_model6.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_max_med, degree=3, raw=TRUE), data = stor_med_mad5, weights = count)
summary(med_model6.1)
residuals6.1 <- residuals(med_model6.1)


med_model6.2 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_max_med,cen_slope_med,degree=3, raw=TRUE), data = stor_med_mad5, weights = count)
summary(med_model6.2)
residuals6.2 <- residuals(med_model6.2)

anova(med_model6.1, med_model6.2)

stor_med_mad5 %>% 
  ggplot() +
  geom_point(aes(x = cen_slope_med, y = residuals6.1)) +
  geom_smooth(aes(x = cen_slope_med, y = residuals6.1))



#Slope_cut, max_curve, plan_curve
med_model6.21 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_max_med, cen_slope_med, degree=3, raw=TRUE), data = stor_med_mad6, weights = count)
summary(med_model6.21)
residuals6.21 <- residuals(med_model6.21)

med_model6.3 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_max_med, cen_slope_med, cen_plan_med, degree=3, raw=TRUE), data = stor_med_mad6, weights = count)
summary(med_model6.3)
residuals6.3 <- residuals(med_model6.3)

anova(med_model6.21, med_model6.3)

stor_med_mad6 %>% 
  ggplot() +
  geom_point(aes(x = cen_plan_med, y = residuals6.21)) +
  geom_smooth(aes(x = cen_plan_med, y = residuals6.21))

#Comparing Models



stor_med_mad5$residuals6.2 <- residuals(med_model6.2)


med_model3.21 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, cen_plan_med, degree=3, raw=TRUE), data = stor_med_mad6, weights = count)
summary(med_model3.21)
stor_med_mad6$residuals3.21 <- residuals(med_model3.21)

med_model6.31 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, cen_max_med, degree=3, raw=TRUE), data = stor_med_mad6, weights = count)
summary(med_model6.31)
stor_med_mad6$residuals6.31 <- residuals(med_model6.31)

anova(med_model3.21,med_model6.31)

sqrt(mean(stor_med_mad6$residuals3.21^2)) #2.08
sqrt(mean(stor_med_mad6$residuals6.31^2)) #1.98: this is better in terms of RMSE & R^squared

```

#Residuals
```{r}
# Obtain the residuals
residuals <- residuals(med_model2)


#Will use this dataset to conduct some EDA 
stor_med_mad$residuals <- residuals



#predicting for Med for every point
pred_med_elev_diff <- predict(med_model2, newdata = stor_no_outliers)

stor_no_outliers$pred_med_elev_diff <- pred_med_elev_diff
```



#Z-Score
```{r}
stor_z <- stor_no_outliers %>% 
  mutate(z_score = (elev_diff - pred_med_elev_diff) / predicted_nmad)

stor_z %>% 
  select(z_score)

# PLots 

    #raster plots
stor_z_raster <- rasterFromXYZ(stor_z) %>% 
  ungroup() %>% 
  dplyr::select(x,y,z_score)

   # geom plots 
stor_z %>% 
  group_by(elev_cut) %>% 
  summarize(med_z_score = median(z_score)) %>% 
  ungroup() %>% 
  ggplot(aes(x = elev_cut, y = med_z_score)) +
  geom_point()

stor_z %>% 
  ggplot(aes(x = elev_cut, y = z_score)) +
  geom_boxplot() +
  scale_y_continuous(limits = c(-5, 5))

stor_z %>% 
  ggplot(aes(x = aspect_cut, y = z_score)) +
  geom_boxplot() +
  scale_y_continuous(limits = c(-50, 50))
```

```{r}
stor_z <- write_csv(stor_z, "data/stor_z.csv")
stor_residuals <- write_csv(stor_med_mad, "data/stor_residuals.csv")
```


#Residual Plots EDA
```{r}
#Plot the residuals against the fitted values
plot(med_model2,1)

plot(fitted(med_model2), residuals,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs. Fitted Values")

stor_med_mad %>% 
  ggplot(aes(elev_cut, residuals))+
  geom_boxplot()

stor_med_mad %>% 
  ggplot(aes(slope_cut, residuals))+
  geom_boxplot()

stor_med_mad %>% 
  ggplot(aes(aspect_cut, residuals))+
  geom_boxplot()

stor_med_mad %>% 
  ggplot(aes(cen_elev_med, residuals))+
  geom_point()

stor_med_mad %>% 
  ggplot(aes(cen_slope_med, residuals))+
  geom_point()

stor_med_mad %>% 
  ggplot(aes(cen_aspect_med, residuals))+
  geom_point()


```
 
#Residual Model
```{r}
#Build a model w/ only cen_elev_med, then find residual, then make a model off of the residual, then residual vs slope; Same as the interpolation model, so use center median of each variable for plotting
res_model1 <- lm(residuals ~ 0 + elev_cut + slope_cut + aspect_cut, data = stor_med_mad)
summary(res_model1)

res_model1.1 <- lm(residuals ~ 0 + cen_elev_med + cen_slope_med + cen_aspect_med, data = stor_med_mad)
summary(res_model1.1)

# res_model2 <- lm(residuals ~ polym(cen_elev_med, cen_slope_med, cen_aspect_med ,degree=2, raw=TRUE), data = stor_med_mad, weights = count)
# 
# res_model2
```

