---
title: "Coreg_eda"
output: html_document
date: '2023-06-30'
editor_options: 
  chunk_output_type: console
---
```{r, message = FALSE}
library(tidyverse)
library(raster)
library(terra)
library(viridis)
library(RColorBrewer)
library(ggplot2)
library(ggplotify)
rwb <- colorRampPalette(c("red", "white", "blue"))(50)

stor_no_outliers <- read_csv("data/stor_no_outliers.csv") #glacier only w/ outliers removed
```

A more cleaner version of CV_Rolstad_MedianModel.rmd. This is only focus on the median model and not the nmad model. For that, please refer to nmad_model.rmd or Meraf_namd.rmd

Once all the exploration and model comparing were done, we came to the conclusion to only use elevation as the only variable to make a model to predict median. Although the other variables such as slope and other curvatures are significant predictors, elevation by itself with an r^2 of 90+%. 

After checking this out, move to Z_score.rmd

#Elevation & Curve to Median
```{r}

stor_no_out_plot <- stor_no_outliers %>% 
  group_by(elev_cut, plan_curve_bin, prof_curve_bin,max_curve_bin,slope_cut) %>% 
  mutate(med = median(elev_diff)) %>% 
  mutate(upper = med + mad(elev_diff)*2) %>% 
  mutate(lower = med - mad(elev_diff)*2) %>% 
  mutate(mad = mad(elev_diff)) 


#Elevation
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = med))

#Curve plan
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = plan_curve_bin, y = med))

#Curve prof
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = prof_curve_bin, y = med))

#Curve max
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = max_curve_bin, y = med))

#Curve max
stor_no_out_plot %>% 
  ggplot() +
  geom_boxplot(aes(x = northness_cut, y = med)) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )

#Curve max
stor_no_out_plot %>%
  ggplot() +
  geom_boxplot(aes(x = eastness_cut, y = med)) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )
```


#EDA
```{r}
stor_no_out <- stor_no_outliers %>% 
  mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
        upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
        med_elev_cut = (lower_elev + upper_elev) / 2,
        cen_elev_med = med_elev_cut - median(med_elev_cut)) 
```


# aspect (northness and eastness)
```{r}
stor_no_out %>% 
  group_by(northness_cut) %>% 
  summarise(med_elev_diff = median(elev_diff),
            count = n())

stor_no_out %>% 
  group_by(eastness_cut) %>% 
  summarize(med_elev_diff = median(elev_diff),
            count = n()) 
```


#MED Data
```{r}
stor_med_elev <- stor_no_out %>% 
  group_by(elev_cut) %>% 
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff),
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut))
 

stor_med_elev_slope <- stor_no_out %>% 
  group_by(elev_cut, slope_cut) %>% 
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff),
         med_slope = median(slope),
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
  mutate(cen_slope_med = med_slope - median(med_slope))

sapply(lapply(stor_med_elev_slope, unique), length)

stor_med_elev_north <- stor_no_out %>% 
  group_by(elev_cut, northness_cut) %>% 
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff),
         med_north = median(northness),
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
  mutate(cen_north_med = med_north - median(med_north))

sapply(lapply(stor_med_elev_north, unique), length)

stor_med_elev_east <- stor_no_out %>% 
  group_by(elev_cut, eastness_cut) %>% 
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff),
         med_east = median(eastness),
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
  mutate(cen_east_med = med_east - median(med_east))

sapply(lapply(stor_med_elev_east, unique), length)


stor_med_elev_slope_north <- stor_no_out %>% 
  group_by(elev_cut, slope_cut, northness_cut) %>% 
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff),
         med_north = median(northness),
         med_slope = median(slope),
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
  mutate(cen_north_med = med_north - median(med_north)) %>% 
  mutate(cen_slope_med = med_slope - median(med_slope))

sapply(lapply(stor_med_elev_slope_north, unique), length)

stor_med_elev_slope_east <- stor_no_out %>% 
  group_by(elev_cut, slope_cut, eastness_cut) %>% 
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff),
         med_east = median(eastness),
         med_slope = median(slope),
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
  mutate(cen_east_med = med_east - median(med_east)) %>% 
  mutate(cen_slope_med = med_slope - median(med_slope))

stor_med_all <- stor_no_out %>% 
  group_by(elev_cut, slope_cut, eastness_cut, northness_cut) %>% 
  #Do two group by at a time, and switch it out to figure out which variable are important and should be included
  summarise(med_elev_diff = median(elev_diff), 
         mad_elev_diff = mad(elev_diff),
         med_east = median(eastness),
         med_slope = median(slope),
         med_north = median(northness),
         count = n()) %>% 
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) %>% 
  mutate(cen_east_med = med_east - median(med_east)) %>% 
  mutate(cen_slope_med = med_slope - median(med_slope)) %>% 
  mutate(cen_north_med = med_north - median(med_north))

```


#MED Models
```{r}
#Cen_elev_med vs cen_slope_med
med_model1 <- lm(med_elev_diff ~ polym(cen_elev_med, degree=3, raw=TRUE), data = stor_med_elev_slope, weights = count)
summary(med_model1)
stor_med_elev_slope$residuals1 <- residuals(med_model1)

stor_med_elev_slope %>% 
  ggplot() +
  geom_point(aes(x = cen_slope_med, y = residuals1)) +
  geom_smooth(aes(x = cen_slope_med, y = residuals1))

med_model1.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, degree=3, raw=TRUE), data = stor_med_elev_slope, weights = count)
summary(med_model1.1)
stor_med_elev_slope$residuals1.1 <- residuals(med_model1.1)

#Partial F-test (Nested Models)
anova(med_model1, med_model1.1) #small p-value, which means add to the model
```


```{r}
#Cen Elev med vs Northness
med_model2 <- lm(med_elev_diff ~ polym(cen_elev_med, degree=3, raw=TRUE), data = stor_med_elev_north, weights = count)
summary(med_model2)
stor_med_elev_north$residuals2 <- residuals(med_model2)

stor_med_elev_north %>% 
  ggplot() +
  geom_point(aes(x = cen_north_med, y = residuals2)) +
  geom_smooth(aes(x = cen_north_med, y = residuals2))

med_model2.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_north_med, degree=3, raw=TRUE), data = stor_med_elev_north, weights = count)
summary(med_model2.1)
stor_med_elev_north$residuals2.1 <- residuals(med_model2.1)

#Partial F-test (Nested Models)
anova(med_model2, med_model2.1) #small p-value, which means add to the model
```


```{r}
#Cen_elev_med vs Eastness
med_model3 <- lm(med_elev_diff ~ polym(cen_elev_med, degree=3, raw=TRUE), data = stor_med_elev_east, weights = count)
summary(med_model3)
stor_med_elev_east$residuals3 <- residuals(med_model3)

stor_med_elev_east %>% 
  ggplot() +
  geom_point(aes(x = cen_east_med, y = residuals3)) +
  geom_smooth(aes(x = cen_east_med, y = residuals3))

med_model3.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_east_med, degree=3, raw=TRUE), data = stor_med_elev_east, weights = count)
summary(med_model3.1)
stor_med_elev_east$residuals3.1 <- residuals(med_model3.1)

#Partial F-test (Nested Models)
anova(med_model3, med_model3.1) #small p-value, which means add to the model
```


```{r}
#Cen_Elev_med & Cen_slope_med vs Northness
med_model4 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, degree=3, raw=TRUE), data = stor_med_elev_slope_north, weights = count)
summary(med_model4)
stor_med_elev_slope_north$residuals4 <- residuals(med_model4)

stor_med_elev_slope_north %>% 
  ggplot() +
  geom_point(aes(x = cen_north_med, y = residuals4)) +
  geom_smooth(aes(x = cen_north_med, y = residuals4))

med_model4.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, cen_north_med, degree=3, raw=TRUE), data = stor_med_elev_slope_north, weights = count)
summary(med_model4.1)
stor_med_elev_slope_north$residuals4.1 <- residuals(med_model4.1)

#Partial F-test (Nested Models)
anova(med_model4, med_model4.1) #small p-value, which means add to the model
```


```{r}
#Cen_elev_med, cen_slope_med vs Eastness

med_model5 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med ,degree=3, raw=TRUE), data = stor_med_elev_slope_east, weights = count)
summary(med_model5)
stor_med_elev_slope_east$residuals5 <- residuals(med_model5)

stor_med_elev_slope_east %>% 
  ggplot() +
  geom_point(aes(x = cen_east_med, y = residuals5)) +
  geom_smooth(aes(x = cen_east_med, y = residuals5))

med_model5.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, cen_east_med, degree=3, raw=TRUE), data = stor_med_elev_slope_east, weights = count)
summary(med_model5.1)
stor_med_elev_slope_east$residuals5.1 <- residuals(med_model5.1)

#Partial F-test (Nested Models)
anova(med_model5, med_model5.1) #small p-value, which means add to the model

```

```{r}
#cen_elev_med, cen_slope_med, cen_east_med vs Northness

med_model6 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, cen_east_med ,degree=3, raw=TRUE), data = stor_med_all, weights = count)
summary(med_model6)
stor_med_all$residuals6 <- residuals(med_model6)

stor_med_all%>% 
  ggplot() +
  geom_point(aes(x = cen_north_med, y = residuals6)) +
  geom_smooth(aes(x = cen_north_med, y = residuals6))

med_model6.1 <- lm(med_elev_diff ~ polym(cen_elev_med, cen_slope_med, cen_east_med, cen_north_med ,degree=3, raw=TRUE), data = stor_med_all, weights = count)
summary(med_model6.1)
stor_med_all$residuals6.1 <- residuals(med_model6.1)

#Partial F-test (Nested Models)
anova(med_model6, med_model6.1) #small p-value, which means add to the model

```






