---
title: "Open and View Storglaciaren"
author: "Laura Boehm Vock"
date: "2023-05-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, message = FALSE}
library(tidyverse)
library(raster)
library(terra)
library(viridis)
library(RColorBrewer)
rwb <- colorRampPalette(c("red", "white", "blue"))(50)
```

```{r}
setwd("~/curi_2023_storglaciaren")

# 2015 data
lidar_2m <- raster("data/Tarfala_Lantmatriet_2015AUG21_2M.tif")

# 2022 data
uav_2m_utm <- raster("data/storglaciaren_2022_sept_uav_2m_SWEREF99_TM.tif")

# glacier outline in 2022
storglacier <- shapefile("data/Storglaciaren_Extents_2022/Storglaciaren_Extents_2022SEP.shp")



lidar_2m
uav_2m_utm
storglacier

plot(lidar_2m, col= magma(64), main = "2015" )
plot(storglacier, add = TRUE)

plot(uav_2m_utm, col = magma(64), main = "2022")
plot(storglacier, add = TRUE)


```

In order to take the difference, we need to be sure the two grids align with each other. 
This process is called *resampling*. 
Read about it here: 
https://desktop.arcgis.com/en/arcmap/latest/extensions/spatial-analyst/performing-analysis/cell-size-and-resampling-in-analysis.htm


```{r}
diff <- resample(uav_2m_utm, lidar_2m) - lidar_2m
```

```{r, eval = FALSE}
# See help menu for resample function
?resample
```


Plot of differences: (Note: try plotting with and without the trim function)
```{r}
plot(trim(diff), col = rwb)
plot(storglacier, add = TRUE)
```

As you can see above, the differences are around 30m, even off the glacier, where there should be no change. 
There is some sort of issue with calibration that we will deal with more in depth later.


The "mask" function will mask out the glacier. Using cellStats, we then calculate the median difference in elevation in the "off glacier" area. 
```{r}
mask(diff, storglacier, inverse = TRUE) %>% cellStats(stat = median)
```


```{r}
diff_adj <- diff - 31.28

plot(trim(diff_adj), col = rwb, zlim = c(-15, 15))
plot(storglacier, add = TRUE)
```


We can look at the slope and aspect as well. 
By default these are both measured in radians
A bit more about aspect here https://gisgeography.com/aspect-map/
and slope here https://pro.arcgis.com/en/pro-app/latest/tool-reference/spatial-analyst/how-slope-works.htm

Together, we can use slope and aspect to draw a "hillShade" which helps us to visualize the topography.
```{r}
stor_terrain <- terrain(lidar_2m, c("slope", "aspect"))
stor_hillShade <- hillShade(stor_terrain$slope, stor_terrain$aspect)

# in this plot I convert slope from radians to degrees
plot(stor_terrain$slope*180/pi, col = colorRampPalette(c("white", "red"))(64), main = "slope (deg)")
plot(storglacier, add = TRUE)

plot(stor_hillShade, col = grey((1:64)/64), main = "Hillshade")
plot(storglacier, add = TRUE)


plot(stor_hillShade, col = grey((1:64)/64), main = "Hillshade +Elevation", legend = FALSE)
plot(lidar_2m, col = magma(13), alpha = 0.4, add =TRUE)
plot(storglacier, add=TRUE)
```


```{r}
names(diff) <- "diff"
names(diff_adj) <- "diff_adj"
names(lidar_2m) <- "elev"

#This will create a glacier indicator
blank <- lidar_2m
values(blank) <- 1
stor_mask <- mask(blank, storglacier) 
names(stor_mask) <- "is.glacier"


stor_stack <- stack(lidar_2m, diff, diff_adj, stor_terrain, stor_mask)

stor_df <- stor_stack %>% 
  as.data.frame() %>%
  as_tibble() %>%
  mutate(is.glacier = ifelse(is.na(is.glacier), 0, 1))


stor_df

summary(stor_df)
```


You can now look at summary stats and plots to see the relationship between diff_adj and elevation, slope, and aspect on and off glacier!!!


```{r}
#EDA
library(mosaic) 
library(tigerstats)
favstats(~slope, data = stor_df)
favstats(~aspect, data = stor_df)
favstats(~diff_adj, data = stor_df)

stor_df <- stor_df%>%
  mutate(is.glacier = as.factor(is.glacier))%>%
  drop_na()

cor_matrix <- stor_df %>%
 dplyr::select(elev,diff,diff_adj,slope,aspect) %>%
  cor()


ggplot(stor_df, aes(elev, diff_adj))+
  geom_point()



ggplot(stor_df, aes(elev, diff_adj))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~is.glacier)

ggplot(stor_df, aes(aspect, diff_adj, color = is.glacier))+
  geom_point()

ggplot(stor_df, aes(aspect, diff_adj, color = is.glacier))+
  geom_point()+
  geom_smooth(method = "lm")

ggplot(stor_df, aes(is.glacier, diff_adj))+
  geom_boxplot()

ggplot(stor_df, aes(aspect,diff_adj, color = is.glacier))+
  geom_violin()

ggplot(stor_df, aes(elev,diff_adj))+
  geom_violin()

ggplot(data = stor_df, mapping = aes(x = diff_adj, colour = is.glacier)) + 
  geom_density()

ggplot(data = stor_df, mapping = aes(x = diff_adj)) +
  geom_histogram(bins = 8) +
  facet_grid(is.glacier ~ .)

ggplot(data = stor_df, mapping = aes(x = diff_adj, fill= is.glacier))+
  geom_histogram()

ggplot(stor_df, aes(elev, diff_adj, color = is.glacier))+
  geom_point()+
  geom_smooth(method=lm,se=FALSE, color = "black")

# outliers 

mad <- mad(stor_df$diff_adj)



stor_df %>%
  mutate(mad = mad(diff_adj))%>%
  filter(mad > 2 )%>%
  ggplot(aes(x =elev, y = diff_adj, color = diff_adj))+
  geom_point()+
  geom_smooth()

summary(diff_adj)
 -1.166367 + mad(stor_df$diff_adj)*2
-1.166367 - mad(stor_df$diff_adj)*2

stor_df %>% 
  filter(diff_adj >= 2.926436 | diff_adj <= -5.25917) %>%
  ggplot() +
  geom_point(aes(x = elev, y = diff_adj)) + 
  facet_grid(.~is.glacier)
  
#raster 
 stor <- stor_df %>% 
   filter(diff_adj >= 2.926436 | diff_adj <= -5.25917) %>%
     dplyr::select(x,y,diff_adj)
 stor <- rasterFromXYZ(stor)
 plot(stor)
 plot(uav_2m_utm, col = magma(64), main = "2022")
 plot(stor, add = TRUE)

 plot(lidar_2m, col= magma(64), main = "2015" )
 plot(stor, add = TRUE)

 #cut off 
 
 stor_df %>% 
  mutate(asp_cut = cut_width(aspect, width= 10)) %>% 
  group_by(asp_cut) %>%
  mutate(med = median(diff_adj),
         upper = med + mad((diff_adj)*2),
         lower = med - mad((diff_adj)*2)) %>% 
  filter(diff_adj >= upper | diff_adj <= lower) %>%
   ggplot(aes(x = asp_cut, y = diff_adj, color = is.glacier))+
   geom_violin()

```

# playing around with raster

```{r}
ras_no_outliers <- stor_na %>% 
  mutate(upper = median(diff_adj) + 2 * mad(diff_adj),
         lower = median(diff_adj) - 2 * mad(diff_adj)) %>% 
  filter(diff_adj < upper, diff_adj > lower) %>% 
     dplyr::select(x,y,diff_adj)
ras_no_outliers <- rasterFromXYZ(ras_no_outliers)

plot(ras_no_outliers)

ras_outliers <- stor_na %>% 
  mutate(upper = median(diff_adj) + 2 * mad(diff_adj),
         lower = median(diff_adj) - 2 * mad(diff_adj)) %>% 
  filter(diff_adj > upper | diff_adj < lower) %>% 
     dplyr::select(x,y,diff_adj)
ras_outliers <- rasterFromXYZ(ras_outliers)

plot(ras_outliers)

plot(ras_no_outliers)


stor_na %>% 
  mutate(upper = median(diff_adj) + 2 * mad(diff_adj),
         lower = median(diff_adj) - 2 * mad(diff_adj)) %>% 
  filter(diff_adj <= upper, diff_adj >= lower) %>% 
  ggplot(aes(x, y, fill = diff_adj)) +
  geom_raster()





```




```{r}
library(stats)


stor_df <- stor_df%>%
  mutate(is.glacier = as.factor(is.glacier))%>%
  drop_na()


data <- stor_df$diff_adj

median_data <- median(stor_df$diff_adj)
mad_data <- mad(stor_df$diff_adj)

threshold <- 2

outliers <- stor_df$diff_adj[abs(stor_df$diff_adj - median_data) > threshold * mad_data]

outliers_df <- data.frame(
  x = outliers,
  y = stor_df$elev
)

ggplot()+
  geom_point(data=outliers_df, aes(x=el, y=diff_adj))
  


#outliers %>%
  #ggplot( ggplot(aes(x =elev, y = diff_adj, color = diff_adj))+
  #geom_point()+
 # geom_smooth())
  
```



 