---
title: "Open and View Storglaciaren"
author: "Laura Boehm Vock"
date: "2023-05-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Most of all the plots here come from CV_open_data_EDA.rmd. There are a few that do not. And for those, you only need to run stor_na to run them. If you want to look for that chunk, look for the section title: New Plot...

```{r, message = FALSE}
library(tidyverse)
library(raster)
library(terra)
library(viridis)
library(RColorBrewer)
library(sf)
library(tibble)
rwb <- colorRampPalette(c("red", "white", "blue"))(50)
```

```{r}
setwd("~/curi_2023_storglaciaren")

# 2015 data
lidar_2m <- raster("data/Tarfala_Lantmatriet_2015AUG21_2M.tif")

# 2022 data
uav_2m_utm <- raster("data/storglaciaren_2022_sept_uav_2m_SWEREF99_TM.tif")

# glacier outline in 2022
storglacier <- shapefile("data/Storglaciaren_Extents_2022/Storglaciaren_Extents_2022SEP.shp")



lidar_2m
uav_2m_utm
storglacier

plot(lidar_2m, col= magma(64), main = "2015" )
plot(storglacier, add = TRUE)

plot(uav_2m_utm, col = magma(64), main = "2022")
plot(storglacier, add = TRUE)


```

In order to take the difference, we need to be sure the two grids align with each other. 
This process is called *resampling*. 
Read about it here: 
https://desktop.arcgis.com/en/arcmap/latest/extensions/spatial-analyst/performing-analysis/cell-size-and-resampling-in-analysis.htm


```{r}
diff <- resample(uav_2m_utm, lidar_2m) - lidar_2m
```

```{r, eval = FALSE}
# See help menu for resample function
?resample
```


Plot of differences: (Note: try plotting with and without the trim function)
```{r}
plot(trim(diff), col = rwb)
plot(storglacier, add = TRUE)
```

As you can see above, the differences are around 30m, even off the glacier, where there should be no change. 
There is some sort of issue with calibration that we will deal with more in depth later.


The "mask" function will mask out the glacier. Using cellStats, we then calculate the median difference in elevation in the "off glacier" area. 
```{r}
mask(diff, storglacier, inverse = TRUE) %>% cellStats(stat = median)
```


```{r}
diff_adj <- diff - 31.28

plot(trim(diff_adj), col = rwb, zlim = c(-15, 15))
plot(storglacier, add = TRUE)
```


We can look at the slope and aspect as well. 
By default these are both measured in radians
A bit more about aspect here https://gisgeography.com/aspect-map/
and slope here https://pro.arcgis.com/en/pro-app/latest/tool-reference/spatial-analyst/how-slope-works.htm

Together, we can use slope and aspect to draw a "hillShade" which helps us to visualize the topography.
```{r}
stor_terrain <- terrain(lidar_2m, c("slope", "aspect"))
stor_hillShade <- hillShade(stor_terrain$slope, stor_terrain$aspect)

# in this plot I convert slope from radians to degrees
plot(stor_terrain$slope*180/pi, col = colorRampPalette(c("white", "red"))(64), main = "slope (deg)")
plot(storglacier, add = TRUE)

plot(stor_hillShade, col = grey((1:64)/64), main = "Hillshade")
plot(storglacier, add = TRUE)


plot(stor_hillShade, col = grey((1:64)/64), main = "Hillshade +Elevation", legend = FALSE)
plot(lidar_2m, col = magma(13), alpha = 0.4, add =TRUE)
plot(storglacier, add=TRUE)
```


```{r}
names(diff) <- "diff"
names(diff_adj) <- "diff_adj"
names(lidar_2m) <- "elev"

#This will create a glacier indicator
blank <- lidar_2m
values(blank) <- 1
stor_mask <- mask(blank, storglacier) 
names(stor_mask) <- "is.glacier"


stor_stack <- stack(lidar_2m, diff, diff_adj, stor_terrain, stor_mask)

stor_df <- stor_stack %>% 
  as.data.frame(xy = TRUE) %>%
  as_tibble() %>%
  mutate(is.glacier = ifelse(is.na(is.glacier), 0, 1))


stor_df

summary(stor_df)


stor_na <- stor_df %>% 
  drop_na() %>% 
  mutate(is.glacier = as.factor(is.glacier))

stor_na <- stor_df %>% 
  drop_na() %>% 
  mutate(is.glacier = as.factor(is.glacier))

stor_df %>% 
  count(is.na(diff))

stor_df %>% 
  filter(is.glacier == 1) %>% 
  count(is.na(diff_adj))
  
```

You can now look at summary stats and plots to see the relationship between diff_adj and elevation, slope, and aspect on and off glacier!!!

#Updated Version of the One Above (If you checked out CV_open_data_EDA already, then you can skip everything in this rmd)
```{r}
mad(stor_na$diff_adj)*2

summary(stor_na)

stor_na_interval <- stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325)) %>% #cut_width & center should be even
  group_by(elev_cut, is.glacier) %>% 
  summarise(n = n()) %>% 
  print(n = 50)

stor_na_width <- stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325)) %>% #cut_width & center should be even
  group_by(elev_cut, is.glacier) %>% 
  mutate(med = median(diff_adj)) %>% 
  mutate(upper = med + mad(diff_adj)*2) %>% 
  mutate(lower = med - mad(diff_adj)*2) %>% 
  filter(diff_adj >= upper | diff_adj <= lower) 

#All (plotting all the points, which are separated by glacier or not)
stor_na %>% 
  ggplot() +
  geom_point(aes(x = x, y = y, color = diff_adj)) +
  facet_grid(.~is.glacier) +
  scale_color_viridis_c(option = "magma") +
  labs(color = "Elevation Difference") +
  coord_equal()


stor_na_outlier <- stor_na %>% 
  filter(diff_adj >= 2.926436 | diff_adj <= -5.25917) # removing everything that is not an outlier. Again, this is still EDA.

#Straight Med Outlier 
stor_na_outlier %>% 
  ggplot() +
  geom_point(aes(x = x, y = y, color = diff_adj)) +
  facet_grid(.~is.glacier) +
  coord_equal()

#Interval Outlier
stor_na %>% 
  inner_join(bind_outliers, by = "diff_adj") %>% 
  ggplot() +
  geom_point(aes(x = x, y = y, color = diff_adj)) +
  facet_grid(.~is.glacier) +
  coord_equal()

#Interval Outlier group_by(elev_cut, is.glacier)
stor_na_width %>% 
  ggplot() +
  geom_point(aes(x = x, y = y, color = diff_adj)) +
  facet_grid(.~is.glacier) +
  scale_color_viridis_c(option = "magma") +
  labs(color = "Elevation Difference") +
  coord_equal()

stor_width_raster <- rasterFromXYZ(stor_na_width %>% 
  ungroup() %>% 
  dplyr::select(x,y,diff_adj))


#raster 
 plot(stor_width_raster)
 plot(uav_2m_utm, col = magma(64), main = "2022")
 plot(stor_width_raster, add = TRUE, show.legend = FALSE)

 plot(lidar_2m, col= magma(64), main = "2015" )
 plot(stor_width_raster, add = TRUE)
```


```{r}
#Bin Plot
stor_na %>% 
  ggplot() +
  geom_point(aes(x = elev, y = diff_adj)) +
  facet_grid(.~is.glacier) 

stor_na2 <- stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325, dig.lab = 4)) %>% #center = 1340???
  group_by(elev_cut, is.glacier) %>% 
  mutate(med = median(diff_adj)) %>% 
  mutate(upper = med + mad(diff_adj)*2) %>% 
  mutate(lower = med - mad(diff_adj)*2)


stor_na2 %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(aes(x = as.numeric(elev_cut), y = upper, color = "mad*2")) + 
  geom_smooth(aes(x = as.numeric(elev_cut), y = med, color = "med")) +
  facet_grid(.~is.glacier) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )

```

#Update No-outlier plot (3 facets: stable terrain, steep terrain, glacier)
```{r}
stor_na4 <- stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325, dig.lab = 4)) %>%
  mutate(slope_degrees = (as.numeric(slope)*180)/pi) %>% 
  mutate(is.steep = ifelse(slope_degrees > 30, "steep", "stable")) %>% 
  group_by(elev_cut, is.glacier, is.steep) %>% 
  mutate(med = median(diff_adj)) %>% 
  mutate(uppest = med + mad(diff_adj)*3) %>% 
  mutate(lowest = med - mad(diff_adj)*3) %>% 
  filter(diff_adj <= uppest | diff_adj >= lowest) %>% 
  mutate(med = median(diff_adj)) %>% 
  mutate(upper = med + mad(diff_adj)*2) %>% 
  mutate(lower = med - mad(diff_adj)*2)

stor_na4.1 <- stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325, dig.lab = 4)) %>%
  mutate(slope_degrees = (as.numeric(slope)*180)/pi) %>% 
  mutate(is.steep = ifelse(slope_degrees > 30, "steep", "stable")) %>% 
  mutate(is.steep = ifelse(is.glacier == 1, "glacier", is.steep)) %>% 
  group_by(elev_cut, is.glacier, is.steep) %>% 
  mutate(med = median(diff_adj)) %>% 
  mutate(uppest = med + mad(diff_adj)*3) %>% 
  mutate(lowest = med - mad(diff_adj)*3) %>% 
  filter(diff_adj <= uppest | diff_adj >= lowest) %>% 
  mutate(med = median(diff_adj)) %>% 
  mutate(upper = med + mad(diff_adj)*2) %>% 
  mutate(lower = med - mad(diff_adj)*2)


stor_na3.2 <- stor_na4 %>% 
  filter(is.glacier == 0) %>% 
  filter(is.steep == "stable")

stor_na3.3 <- stor_na4 %>% 
  filter(is.glacier == 0) %>% 
  filter(is.steep == "steep")

stor_na3.4 <- stor_na4 %>% 
  filter(is.glacier == 1) 
  
stor_na3 %>% 
  dplyr::select(zscore, is.steep) 

#test
stor_na_no_med <- stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325, dig.lab = 4)) %>%
  mutate(slope_degrees = (as.numeric(slope)*180)/pi) %>% 
  mutate(is.steep = ifelse(slope_degrees > 30, "steep", "stable")) %>% 
  group_by(elev_cut, is.glacier, is.steep) %>% 
  mutate(med = median(diff_adj)) %>% 
  mutate(uppest = med + mad(diff_adj)*3) %>% 
  mutate(lowest = med - mad(diff_adj)*3) %>% 
  filter(diff_adj < uppest | diff_adj > lowest) %>% 
  mutate(med = median(diff_adj)) %>% 
  mutate(upper = med + mad(diff_adj)*2) %>% 
  mutate(lower = med - mad(diff_adj)*2)

stor_na4 %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(aes(x = as.numeric(elev_cut), y = upper, color = "mad*2")) + 
  geom_smooth(aes(x = as.numeric(elev_cut), y = med, color = "med")) +
  facet_grid(is.steep~is.glacier) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        )

stor_na4.1 %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(aes(x = as.numeric(elev_cut), y = upper, color = "mad*2")) + 
  geom_smooth(aes(x = as.numeric(elev_cut), y = med, color = "med")) +
  facet_grid(.~is.steep) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 7)
        ) +
  labs(y = "Elevation Difference",
       x = "Elevation Interval")


#p1 <- 
stor_na3.2 %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(aes(x = as.numeric(elev_cut), y = upper, color = "mad*2")) + 
  geom_smooth(aes(x = as.numeric(elev_cut), y = med, color = "med")) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        ) +
  labs(title = "non-glacier stable terrain")

#test
stor_na_no_med %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(aes(x = as.numeric(elev_cut), y = upper, color = "mad*2")) + 
  geom_smooth(aes(x = as.numeric(elev_cut), y = med, color = "med")) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        ) +
  labs(title = "non-glacier stable terrain")

#p2 <- 
  stor_na3.3 %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(aes(x = as.numeric(elev_cut), y = upper, color = "mad*2")) + 
  geom_smooth(aes(x = as.numeric(elev_cut), y = med, color = "med")) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        ) +
  labs(title = "non-glacier steep terrain")

#p3 <- 
 stor_na3.4 %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(aes(x = as.numeric(elev_cut), y = upper, color = "mad*2")) + 
  geom_smooth(aes(x = as.numeric(elev_cut), y = med, color = "med")) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        ) +
  labs(title = "glacier")
 
 #test_p3 <- 
 stor_na_no_med %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(aes(x = as.numeric(elev_cut), y = upper, color = "mad*2")) + 
  geom_smooth(aes(x = as.numeric(elev_cut), y = med, color = "med")) +
  theme(axis.text.x = element_text(angle = 45),
        text = element_text(size = 8)
        ) +
  labs(title = "glacier")

library(gridExtra)


```



#New Plots that were not in CV_open_data, but were used in our presentation slides. They are beautiful to look at, but still serve as better eda than the CV_open_data_EDA.

```{r}
# Setting up the datasets to run the plots below
stor_no_outliers1 <- stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325, dig.lab = 4)) %>%  mutate(slope_degrees = (as.numeric(slope)*180)/pi) %>% 
  mutate(is.steep = ifelse(slope_degrees > 30, "steep", "stable")) %>% 
  mutate(is.steep = ifelse(is.glacier == 1, "glacier", is.steep)) %>% 
  group_by(elev_cut, is.steep) %>% 
  mutate(med_with_outliers = median(diff_adj), 
         mad_with_outliers = mad(diff_adj)) %>% 
  mutate(uppest = med_with_outliers + mad_with_outliers*3) %>% 
  mutate(lowest = med_with_outliers - mad_with_outliers*3) %>% 
  filter(diff_adj <= uppest, diff_adj >= lowest) %>%
  ungroup() %>%  #### I just added this! Probably
  mutate(aspect_degrees = (as.numeric(aspect)*180)/pi) %>% 
  mutate(aspect_cut = cut_width(aspect_degrees, width = 10, center = 25)) %>% 
  mutate(slope_cut = cut(slope, breaks = c(0, 2.5, 5, 10, 15, 20, 30, 40, 50, 70, 90)))


elev_binned_stats <- stor_no_outliers1 %>% 
  group_by(elev_cut, is.steep) %>%
  summarize(med = median(diff_adj),
            mad = mad(diff_adj)) %>%
  mutate(upper = med + 2*mad, 
         lower = med - 2*mad)

aspect_binned_stats <- stor_no_outliers1 %>% 
  group_by(aspect_cut, is.steep) %>%
  summarize(med = median(diff_adj),
            mad = mad(diff_adj)) %>%
  mutate(upper = med + 2*mad, 
         lower = med - 2*mad)

slope_binned_stats <- stor_no_outliers1 %>% 
  group_by(slope_cut, is.glacier) %>% 
  mutate(slope_cut = labels = c("0-2.5", "2.5-5", "5-10", "10-15", "15-20", "20-30", 
                                    "30-40", "40-50", "50-70", "70-90")) %>% 
  summarize(med = median(diff_adj),
            mad = mad(diff_adj)) %>%
  mutate(upper = med + 2*mad, 
         lower = med - 2*mad)


#Plots

## Elev_cut
stor_no_outliers1 %>% 
  ggplot() +
  geom_boxplot(aes(x = elev_cut, y = diff_adj)) +
  geom_smooth(data = elev_binned_stats, aes(x = as.numeric(elev_cut), y = upper, color = "mad*2"), se = FALSE) + 
    geom_smooth(data = elev_binned_stats, aes(x = as.numeric(elev_cut), y = lower, color = "mad*2"), se= FALSE) + 
  geom_smooth(data = elev_binned_stats, aes(x = as.numeric(elev_cut), y = med, color = "med"), se = FALSE) +
  facet_grid(.~is.steep) +
  theme(axis.text.x = element_text(angle = 45, size = 8),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 15)
        ) +
  labs(y = "Elevation Difference",
       x = "Elevation Interval")


stor_no_outliers1 %>% 
  ggplot() +
  geom_boxplot(aes(x = aspect_cut, y = diff_adj)) +
  geom_smooth(data = aspect_binned_stats, aes(x = as.numeric(aspect_cut), y = upper, color = "mad*2"), se = FALSE) + 
    geom_smooth(data = aspect_binned_stats, aes(x = as.numeric(aspect_cut), y = lower, color = "mad*2"), se= FALSE) + 
  geom_smooth(data = aspect_binned_stats, aes(x = as.numeric(aspect_cut), y = med, color = "med"), se = FALSE) +
  facet_grid(.~is.steep) +
  theme(axis.text.x = element_text(angle = 45, size = 10),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 15)
        ) +
  labs(y = "Elevation Difference",
       x = "Aspect Interval")


stor_no_outliers1 %>%
  ggplot(aes(x,y, fill = diff_adj)) +
  geom_raster() +  
  scale_fill_gradient2(low = "#ca0020", mid = "white", high = "#0571b0",
                       midpoint = 0, limits = c(-15, 15)) + 
  theme_void() + 
  theme(plot.background = element_rect(fill = "lightgrey"))+
  coord_equal() +
  labs(fill = "Elevation Difference")



#Interval Outlier group_by(elev_cut, is.glacier)
stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325, dig.lab = 4)) %>%  
  mutate(slope_degrees = (as.numeric(slope)*180)/pi) %>% 
  mutate(is.steep = ifelse(slope_degrees > 30, "steep", "stable")) %>% 
  mutate(is.steep = ifelse(is.glacier == 1, "glacier", is.steep)) %>% 
  group_by(elev_cut, is.steep) %>% 
  mutate(med_with_outliers = median(diff_adj), 
         mad_with_outliers = mad(diff_adj)) %>% 
  mutate(uppest = med_with_outliers + mad_with_outliers*3) %>% 
  mutate(lowest = med_with_outliers - mad_with_outliers*3) %>% 
  filter(diff_adj >= uppest | diff_adj <= lowest) %>%
  ggplot() +
  geom_point(aes(x = x, y = y, color = diff_adj)) +
  facet_grid(.~is.glacier) +
  scale_color_viridis_c(option = "magma") +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 15)) +
  labs(color = "Elevation Difference") +
  coord_equal()

stor_na %>% 
  mutate(elev_cut = cut_width(elev, width = 50, center = 1325, dig.lab = 4)) %>%  
  mutate(slope_degrees = (as.numeric(slope)*180)/pi) %>% 
  mutate(is.steep = ifelse(slope_degrees > 30, "steep", "stable")) %>% 
  mutate(is.steep = ifelse(is.glacier == 1, "glacier", is.steep)) %>% 
  group_by(elev_cut, is.steep) %>% 
  mutate(med_with_outliers = median(diff_adj), 
         mad_with_outliers = mad(diff_adj)) %>% 
  mutate(uppest = med_with_outliers + mad_with_outliers*3) %>% 
  mutate(lowest = med_with_outliers - mad_with_outliers*3) %>% 
  filter(diff_adj >= uppest | diff_adj <= lowest) %>% 
  ggplot(aes(x,y, fill = diff_adj)) +
  geom_raster() +  
  scale_fill_gradient2(low = "#ca0020", mid = "white", high = "#0571b0",
                       midpoint = 0, limits = c(-15, 15)) + 
  theme_void() + 
  theme(plot.background = element_rect(fill = "lightgrey"))+
  coord_equal() +
  labs(fill = "Elevation Difference")
```



#coregistered_interpolation.rmd is the next thing to check 
