---
title: "Coreg_eda"
output: html_document
date: '2023-06-30'
editor_options: 
  chunk_output_type: console
---
```{r, message = FALSE}
library(tidyverse)
library(raster)
library(terra)
library(viridis)
library(dplyr)
library(RColorBrewer)
rwb <- colorRampPalette(c("red", "white", "blue"))(50)

stor_cor_glacier <- read_csv("data/stor_cor_glacier.csv") 

```

#REFER to CV_Rolstad_MedModel2.rmd for model testing and more information. 

The purpose of this rmd is to identify the z-sore, which will be used to create a variogram that will be essential for error of propogation 


## Set-UP
```{r}
#Median
## Finding the median of the elevation by elevation cuts
### Note that the centers are needed as we are making a polynomial model

stor_med_elev <- stor_cor_glacier %>% 
  group_by(elev_cut) %>% 
  #We only used elevation because its R^2 was already within the 90s. Although the other variables were significant, they did not effect the R^2 too much
  summarise(med_elev_diff = median(elev_diff_no_outliers, na.rm = TRUE), # filled_elev_diff instead??
         count = n()) %>% #counts will play a role in making the model
  ungroup() %>% 
     mutate(lower_elev = as.numeric(gsub("\\((.*),.*\\]", "\\1", elev_cut)),
            upper_elev = as.numeric(gsub(".*,(.*)\\]", "\\1", elev_cut)),
          med_elev_cut = (lower_elev + upper_elev) / 2,
          cen_elev_med = med_elev_cut - median(med_elev_cut)) #1475 is the median(med_elev_cut)

sapply(lapply(stor_med_elev, unique), length) #helps identify the different unique levels/values in the data of each col



#NMAD
#data frame that is going to be used for modeling 
### Unsure, come back to
stor_slope_maxcurv <- stor_cor_glacier %>%
  mutate(max_curv_5 = ifelse(max_curv < 5, max_curv, NA)) %>% 
  # combined the two largest max curve bins and the three largest slope bins to keep the count consistent
 #mutate(max_curve_bin = fct_collapse(max_curve_bin, "(2.36,269]" = c("(2.36,3.26]", "(3.26,269]"))) %>% 
  mutate(slope_cut = fct_collapse(slope_cut, "(26.4,87.2]" = c("(26.4,30.4]","(30.4,34]", "(34,37.2]", "(37.2,42]", "(42,50.8]", "(50.8,87.2]"))) %>% 
  group_by(slope_cut, max_curve_bin) %>% 
  summarise(med_elev_diff = median(elev_diff_no_outliers, na.rm = TRUE), 
         mad_elev_diff = mad(elev_diff_no_outliers, na.rm = TRUE),
         med_slope = median(slope_degrees),
         med_max = median(max_curv_5, na.rm = TRUE),
         ct = n()) %>% 
  ungroup() %>% 
  mutate(cen_max = med_max - median(med_max)) %>% #center of max curvature 
  mutate(cen_slope = med_slope - median(med_slope)) #center of slope 


sapply(lapply(stor_slope_maxcurv, unique), length)

stor_slope_maxcurv %>% 
  slice_sample(n = 5000) %>%
  ggplot(aes(x = max_curve_bin, y = slope_cut, size = ct)) +
  geom_point()


# stor_slope_maxcurv %>% 
#   group_by(max_curve_bin) %>% 
#   summarise(n = n()) %>% 
#   print(n = Inf) 





```

#Dataset to predict on
```{r}
#data frame used to predict nmad and median
## For the data fram we will predict on, we used their actual variables instead of cuts to help center it
stor_cor_glacier <- stor_cor_glacier %>% 
  mutate(cen_elev_med = elev - median(stor_med_elev$med_elev_cut),  #1450
        cen_slope = slope_degrees - median(stor_slope_maxcurv$med_slope), #21.01295
        cen_max = max_curv - median(stor_slope_maxcurv$med_max, na.rm = TRUE)) #1.026125
```


#Med_Model & predict
```{r}
med_model <- lm(med_elev_diff ~ polym(cen_elev_med, degree=3, raw=TRUE), data = stor_med_elev, weights = count) #we choose this model despite having significance in the other ones b/c elev already accounted for 90% of the data (R^2 =90+%)
summary(med_model)
#stor_med_elev$med_residuals <- c(NA, residuals(med_model)) #???


#predicting for Med for every point
stor_cor_glacier$pred_med_elev_diff <- predict(med_model, newdata = stor_cor_glacier)
```

#NMAD_Model & predict
```{r}
#models 

model1_cen <- lm(mad_elev_diff~cen_slope + cen_max + cen_slope:cen_max, 
                    data = stor_slope_maxcurv[stor_slope_maxcurv$cen_slope<11.1- median(stor_slope_maxcurv$med_slope),]) # uses center slope (this is what we used to get the z score, we are predicting on stor_no_out)
model2_cen <- lm(mad_elev_diff~cen_max, 
                    data = stor_slope_maxcurv[stor_slope_maxcurv$cen_slope>11.1 - median(stor_slope_maxcurv$med_slope),]) #for center slope our condition is different because med_slope and cen_slope are not the same based on our graphs from the bottom we picked 11.3 by looking at the end of the bin of slope_degree that falls in the area of cen_slope around 0


summary(model1_cen)
summary(model2_cen)


stor_cor_glacier <- stor_cor_glacier %>% 
  # leave max_curv as it is as we dont use it in our model
 # mutate(cen_slope = ifelse(cen_slope > 12, 12, cen_slope))%>%
  mutate(cen_max = ifelse(max_curv > 2.36, 3.26 - median(stor_slope_maxcurv$med_max), cen_max), 
    pred_mad = ifelse(slope_degrees < 11.1, predict(model1_cen, newdata = .), predict(model2_cen, newdata = .))) #instead of using 12 as our condition we switched it to 11.3 because it was the end of the bin at center slope 0/1

```


## Problem 1: There are negative predicted MAD in stor_no_out
```{r}
stor_cor_glacier$pred_mad %>% summary()

stor_cor_glacier %>%
  slice_sample(n = 5000) %>%
  ggplot(aes(cen_slope, pred_mad, col = max_curv)) + 
  geom_point()

```

 
```{r}
stor_slope_maxcurv_cen <- stor_slope_maxcurv %>% #same thing but uses center slope and switched our condition to less than 1 
  mutate(cen_max = ifelse(cen_max > 3, 3, cen_max)) %>%
  #mutate(med_slope = ifelse(med_slope > 30, 30, med_slope), 
  mutate(pred_mad = ifelse(cen_slope <11.1 - median(stor_slope_maxcurv$med_slope), predict(model1_cen, newdata = .), predict(model2_cen, newdata = .))) 

stor_slope_maxcurv_cen %>% 
  ggplot(aes(x = cen_slope, color=max_curve_bin)) +
  geom_point(aes(y = mad_elev_diff)) +
  geom_line(aes(y = pred_mad)) 

stor_slope_maxcurv %>% 
  ggplot(aes(cen_max, mad_elev_diff)) +
  geom_point()


stor_cor_glacier %>%
  slice_sample(n = 5000) %>%
  ggplot(aes(cen_slope, pred_mad, col = max_curv)) + 
  geom_point()

# stor_cor_glacier %>% 
#   filter(pred_mad < 0) %>% 
#   ggplot(aes(cen_slope, cen_max, col = pred_mad)) +
#   geom_point()


stor_cor_glacier %>% 
  slice_sample(n = 5000)%>%
  mutate(mad_elev_diff = mad(elev_diff)) %>% 
  ggplot(aes(x = cen_slope, color=max_curve_bin)) +
  geom_point(aes(y = mad_elev_diff)) +
  geom_line(aes(y = pred_mad))

# stor_slope_maxcurv%>%
#   ggplot(aes(x = fct_reorder(slope_cut, med_slope), y = fct_reorder(max_curve_bin, cen_max), fill = mad_elev_diff))+
#   geom_tile(inherit.aes = TRUE)#trying to replicate the same graph as the hugonnet figure 4

# library(ggplot2)  
# my_color_palette <- scale_fill_viridis_c()
# 
# # Plotting with ggplot
# 
# install.packages("patchwork")
# library(patchwork)
# 
# plot1 <- stor_slope_maxcurv%>%
# ggplot(aes(x = fct_reorder(slope_cut, med_slope), y = fct_reorder(max_curve_bin, cen_max), fill = mad_elev_diff)) +
#   geom_tile() +
#  scale_fill_viridis_c() +
#   labs(title = "Storglaciaren Volume Change",
#        x = "Median Slope",
#        y = "Central Max Curve",
#        fill = "MAD Elevation Difference") +
#   theme_minimal()
# 
# plot2 <- rasterFromXYZ(stor_z)
#   
# stor_z_raster <- rasterFromXYZ(stor_z) %>% 
#   dplyr::select(x,y,z_score)
# 
#  plot(stor_z_raster)

stor_slope_maxcurv %>% 
  ggplot(aes(x = cen_max, y = pred_mad)) +
  geom_point()

stor_slope_maxcurv %>% 
  slice_sample(n = 5000)%>%
  ggplot(aes(x = pred_mad, y = cen_slope, color = cen_max)) +
  geom_point()


```

#Z-score
```{r}
stor_cor_glacier <- stor_cor_glacier %>% 
  mutate(z_score = (elev_diff_no_outliers - pred_med_elev_diff) / pred_mad)

stor_cor_glacier %>% 
  dplyr::select(z_score, elev_diff_no_outliers, pred_med_elev_diff, pred_mad)

summary(stor_cor_glacier$z_score)
sd(stor_cor_glacier$z_score, na.rm = TRUE)
hist(stor_cor_glacier$z_score)
hist(stor_cor_glacier$pred_mad)

```
 

```{r}
stor_cor_glacier %>% 
   slice_sample(n=5000) %>% 
  ggplot(aes(x = cen_slope, y = pred_mad)) +
  geom_point()

stor_cor_glacier %>% 
  slice_sample(n=5000) %>% 
  ggplot(aes(x = max_curv, y = pred_mad)) +
  geom_point()
# cen_slope > 30, do the same thing like the 

stor_cor_glacier %>% 
  slice_sample(n=5000) %>% 
  ggplot(aes(x = pred_mad, y = slope_degrees, color = max_curv)) +
  geom_point()
  

```

After this rmd, please check out Variogram-Rolstad (variogram that is spherical), from the rolstad reading. For the Variogram that follows Hugonnet, then check out nmad_variigram.rmd
 
```{r}
#write_csv(stor_cor_glacier, "data/stor_cor_glacier.csv") 
```

